---
title: "Title of your project"
author: "your name"
date: "todays' date"
output: html_document
bibliography: your-bib-file.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Project summary

This is an R Markdown template for your project. Start with a general summary. Here you should add a few sentences on the general research question. Tell us how your question will be answered and where the data will come from (e.g., field, lab, simulation, or some other source). Any predictions should be mentioned here.


## 2. Data summary

Here you should provide more detailed information on your data. Below you will have your first R code chunk with data wrangling steps and basic summary statistics of the data.

Embed an R code chunk like this:

```{r my data}

summary(cars)

```

## Plots

Provide histograms of distribution or plots of correlations. 

```{r plot of pressure, echo=FALSE}

plot(pressure)

```

## 3. Data Analyses

Here you should tell us what is planned for the analysis and statistical tests. How does that relate to your research question? why did you choose that particular analysis?


Embed an R code chunk with analysis and result figures

```{r analyses}


```


## 4. Results and Conclusions

What did you find? How do you interpret the results?



## 5. References

Finally, you can add citations. To do this you use the syntax

@R-base

Where R-base is the name of the R citation (used an example below). The @ sign tells R to find that particular citation in the .bib file.


A BibTeX database is a plain-text file (with the conventional filename extension .bib) that consists of bibliography entries like this:

---
@Manual{R-base,
  title = {R: A Language and Environment for Statistical Computing},
  author = {{R Core Team}},
  organization = {R Foundation for Statistical Computing},
  address = {Vienna, Austria},
  year = {2016},
  url = {https://www.R-project.org/},
}
---

You may see https://en.wikipedia.org/wiki/BibTeX for possible types of entries and fields in BibTeX.


A bibliography entry starts with @type{, where type may be article, book, manual, and so on.7 Then there is a citation key, like R-base in the above example. To cite an entry, use @key or [@key] (the latter puts the citation in braces), e.g., @R-base is rendered as R Core Team (2019), and [@R-base] generates “(R Core Team 2019)”. 

#### Citation Syntax

Don't forget that citations go inside square brackets and are separated by semicolons. 

---
Blah blah [@smith04; @doe99]

---