---
title: "W2_Exercise"
author: "Reni Kaul"
date: "4/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Practicing commenting and commentary

The first workshop mentioned the value of consistent coding style and was based on a [tidyverse style guide](http://style.tidyverse.org/index.html). 

> Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread.

Good commenting and commentary also make your code easier to read. This is also a key part of reproducible research. A well documented analysis can makes your work accessible to other scientists, including your future self. 

**Directions**

We will practice this skill by modifying the group exercises from the first workshop with comments and commentary. Start by modifying your group's exercise in the space below *## Documenting my group's exercise*, then add comments and commentary to another group's exercise. You should have a copy of your exploration in your email. 

[DIRECTIONS ON HOW TO GET ANOTHER GROUP'S CODE]. Since you were not involved in their exploration, it might be easier to understand their code by running pieces of it at a time. In order to do that we need to load the `tidyverse` package and passenger data. 

When you are finished adding comments and commentary knit the `.Rmd` file into an HTML. Do the code and figures look how you expected? 

How does your documentation compare with other group members? 

```{r load passenger1 data}
# Import the Tidyverse
library(tidyverse)
# Import data and convert survived to yes/no factor for easy plotting (passengers1).  
passengers <- read_csv("data/train.csv")

passengers1 <- passengers %>% 
  mutate(Survived = ifelse(Survived == 0, "No", "Yes"))
```

*Useful information* 

* The `.Rmd` allows us to mix code (code chunks) and natural language. 
* The natural language can be quickly formatted using `R`'s version of a markdown language ([See page 1 for more details on formatting with Rmarkdown](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)). 
* A minimum code chunk starts with *```{r}* and ends with *```*. You may want to seperate different coding tasks into different code chunks ([See page 2 for more details on the code chunk options](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)). 
* Note that `#` sign within a code chunk starts a comment, but is a section header in Rmarkdown. 

## Documenting my group's exercise


```{r my groups exercise, echo=TRUE}

```


## Documenting another group's exercise


```{r another groups exercise, echo=TRUE}


```

